---
title: "R Datalab"
output:
  html_document: default
  html_notebook: default
---

This is an example of running an R version of [Google Datalab](https://cloud.google.com/datalab/)

Google Datalab is a service that lets you easily interact with your data in the Google Cloud.  This document is an excercise in trying to replicate the same functionality:

* Runs on Google Cloud infrastructure using `googleComputeEngineR` within its own Docker container
* Uses RStudio and its [RMarkdown Notebooks](http://rmarkdown.rstudio.com/r_notebooks.html) to replicate the Jupyter/iPython functionality
* Auto authentication with the Google cloud services to work with BigQuery and Cloud Storage data
* Cross language support of python, SQL and bash via R Notebooks
* Python data analysis libraries [pandas](http://pandas.pydata.org/) and [NumPy](http://www.numpy.org/)
* Visualisation via R libraries such as the [htmlwidgets family](http://gallery.htmlwidgets.org/)
* Installation of [Tensorflow](https://www.tensorflow.org) and RStudio's [tensorflow](https://rstudio.github.io/tensorflow/) package
* Installation of [feather](https://blog.rstudio.org/2016/03/29/feather/) to help R and python share data nicely.

## Setup

```{r, echo=TRUE}
library(googleAuthR)
## this reuses the authentication of the GCE instance we are on
gar_gce_auth()

library(bigQueryR)
## list authenticated projects
myproject <- bqr_list_projects()

library(googleCloudStorageR)
## list Cloud Storage buckets
gcs_list_buckets(myproject$id[[1]])

```

Demo of running python in same document:

```{python, echo=TRUE}
hiss = 'sssssssss'
print "Pythons go %s." % hiss
```

Also works with `SQL` and `bash`

```{bash, echo=TRUE}
pip freeze
```

## Transfer data between R and Python with feather

From the [example intro blogpost](https://blog.rstudio.org/2016/03/29/feather/) for feather:

```{r, echo=TRUE}
library(feather)
df <- mtcars
path <- "my_data.feather"
write_feather(df, path)
```


```{python, echo=TRUE}
import feather
path = 'my_data.feather'
df = feather.read_dataframe(path)
df.head
```

## Tensorflow

### Hello world Python

```{python, echo=TRUE}
from __future__ import print_function

import tensorflow as tf

# Simple hello world using TensorFlow

# Create a Constant op
# The op is added as a node to the default graph.
#
# The value returned by the constructor represents the output
# of the Constant op.
hello = tf.constant('Hello, TensorFlow!')

# Start tf session
sess = tf.Session()

# Run the op
print(sess.run(hello))
```

### Hello world R

```{r, echo=TRUE}
library(tensorflow)
sess = tf$Session()
hello <- tf$constant('Hello, TensorFlow!')
sess$run(hello)
```


## Build details

This was run in a local R session to start up this RStudio instance with the right libraries installed:

```r
library(googleComputeEngineR)

## make an RStudio instance to base upon
vm <- gce_vm(template = "rstudio", 
             name = "r-datalab-build", 
             username = "mark", password = "mark1234", 
             predefined_type = "n1-standard-1")

## once RStudio loaded at the IP, build the Dockerfile below on instance
## this takes a while
docker_build(vm, dockerfolder = get_dockerfolder("cloudDataLabR"), new_image = "r-datalab")


## send to the Container Registry
gce_push_registry(vm, save_name = "datalab-r-image", image_name = "r-datalab")

## Can now launch instances using this image via:
vm2 <- gce_vm(template = "rstudio", 
              name = "r-datalab", 
              predefined_type = "n1-standard-1", 
              dynamic_image = gce_tag_container("datalab-r"), 
              username = "mark", password = "mark1234")

```

The Dockerfile used is below:

```
FROM rocker/hadleyverse
MAINTAINER Mark Edmondson (r@sunholo.com)

# install cron and nano and tensorflow
RUN apt-get update && apt-get install -y \
    cron nano \
    python-pip python-dev
    && pip install cython \
    && pip install numpy \
    && pip install pandas \
    && export TF_BINARY_URL=https://storage.googleapis.com/tensorflow/linux/cpu/tensorflow-0.11.0-cp27-none-linux_x86_64.whl \
    && pip install --upgrade $TF_BINARY_URL \

    && pip install feather-format \
    ## clean up
    && apt-get clean \ 
    && rm -rf /var/lib/apt/lists/ \ 
    && rm -rf /tmp/downloaded_packages/ /tmp/*.rds
    
## Install packages from CRAN
RUN install2.r --error \ 
    -r 'http://cran.rstudio.com' \
    googleAuthR googleAnalyticsR searchConsoleR googleCloudStorageR bigQueryR htmlwidgets feather \
    ## install Github packages
    && Rscript -e "devtools::install_github(c('MarkEdmondson1234/youtubeAnalyticsR', 'MarkEdmondson1234/googleID', 'MarkEdmondson1234/googleAuthR'))" \
    && Rscript -e "devtools::install_github(c('bnosac/cronR'))" \
    && Rscript -e "devtools::install_github(c('rstudio/tensorflow'))" \
    ## clean up
    && rm -rf /tmp/downloaded_packages/ /tmp/*.rds \

```


